---
title: "first markdown file"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```

## Reading the Data

This is data that shows the favorite restaurants of students. Details relevant to each restaurant are included.

```{r}
class <- read_csv("C://users//ra4ct/Desktop//fav_restaurants.csv")
```

## Plotting the Data

```{r}
ggplot(class, aes(cuisine)) +
  geom_bar()
```

This is a graphic showing the popularity of type of cuisine across all of the restaurants. 

```{r}
ggplot(class, aes(yearly_visits, cost_per_person)) +
  geom_point()
```

This is a scatter plot showing the correlation between the avergage cost per person at the restaurant and the amount of times that the restaurant is visited by that person per year. 

```{r}
ggplot(class, aes(location)) +
  geom_bar()
```

This is a bar graph showing the location of each restaurant. There are many restaurants that are only in one location, but many people's favorite restaurants are in the same location.