---
title: "first markdown file"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

## Reading the Data

This is data that shows the *favorite restaurants* of students. Details relevant to each restaurant are included.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
class <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-rosiea26/master/fav_restaurants.csv")
```

## Plotting the Data

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(class, aes(cuisine)) +
  geom_bar()
```

This is a graphic showing the **popularity of type of cuisine** across all of the restaurants. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(class, aes(yearly_visits, cost_per_person)) +
  geom_point()
```

This is a scatter plot showing the **correlation between the avergage cost per person at the restaurant and the amount of times that the restaurant is visited** by that person per year. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(class, aes(location)) +
  geom_bar()
```

This is a bar graph showing the **location of each restaurant**. There are many restaurants that are only in one location, but many people's favorite restaurants are in the same location.